---
title: "Returns of Financial Stocks"
author: ""
date: ""
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  # display code as typed
  tidy=FALSE,     
  # slightly smaller font for code
  size="small")   
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest)    # scrape websites
library(purrr)  
library(lubridate) #to handle dates
library(kableExtra)
library(knitr)
library(dplyr)
library(gapminder)
library(ggplot2)
library(patchwork)
library(scales)
library(tidytext)
```

In the following we perform analyses on Returns of a select number of Financial Stocks utilizing the monthly, yearly and daily returns to gauge behavior of these stocks over a fixed period of time. We obtain our data from the `nyse.csv` dataset which contains 508 stocks listed on the NYSE, their ticker `symbol`, `name`, the IPO year, and the sector and industry the company operates in.

First of all we have to load our data set:

```{r load_nyse_data, message=FALSE, warning=FALSE}

nyse <- read.csv(here::here("data","nyse.csv"))

```

Based on this data set, we create a table and a bar plot that display the number of companies per sector in descending order.

```{r companies_per_sector}

# count the number of companies by sector

companies_by_sector <- nyse %>%
  
  group_by(sector) %>%
  
  count() %>%
  
  arrange(desc(n))

```

```{r}

companies_by_sector %>% 
  
  # rename for better reading
  
  rename(`number of companies` = n) %>% 
  
  # create html table
  
  kbl() %>%
  
  kable_styling()
```

```{r} 
companies_by_sector <- nyse %>%
  
  group_by(sector) %>%
  
  summarise(count = n()) 

#create a bar plot

ggplot(companies_by_sector,
       aes(x = reorder(sector, count), y = count)) + 
  
  geom_col(fill = "#1380A1") + 
  
  theme_bw()+
  
  theme(plot.title = element_text(face = 'bold', size = 15), 
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size= 8)) +

# add label to each bar
  geom_label(aes(label = count), 
             hjust = 1,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             color = 'White',
             size = 4) +
  
  labs(x = "", 
       y = "Number of Companies", 
       title = 'Finance & Consumer companies reign NYSE',
       subtitle = 'NYSE number of companies by sector') +
  
  coord_flip()

```

> From the table and the plot we can infer that Finance and Consumer Services stocks reign NYSE. Transportation and Consumer Durables on the other hand are the least represented industries.

Next, we choose six stocks and their ticker symbols and download the respective data.

```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}

# Notice the cache=TRUE argument in the chunk options. Because getting data is time consuming

# cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

myStocks <- c("AXP","BLK","KO","AMZN","MCD","BP","SPY" ) %>%
  
  tq_get(get  = "stock.prices",
         from = "2011-01-01",
         to   = "2020-08-31") %>%
  
  group_by(symbol)

# examine the resulting data frame

glimpse(myStocks) 

```

Since the financial performance of a stock largely depends on a company's returns, we now calculate daily and monthly returns.

```{r calculate_returns, message=FALSE, warning=FALSE, cache=TRUE}

#calculate daily returns

myStocks_returns_daily <- myStocks %>%
  
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "daily", 
               type       = "log",
               col_rename = "daily_returns",
               cols = c(nested.col))  

#calculate monthly  returns

myStocks_returns_monthly <- myStocks %>%
  
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_returns",
               cols = c(nested.col)) 
```

Now that we have specified monthly and daily returns we can create a table containing the return data for the selected stocks.

```{r summarise_monthly_returns}

myStocks_returns_monthly_summary <- myStocks_returns_monthly %>%
  
  group_by(symbol) %>%
  
  summarise(min = min(monthly_returns), 
            median = median(monthly_returns), 
            max = max(monthly_returns), 
            mean = mean(monthly_returns), 
            sd = sd(monthly_returns))

# calculate summary statistics for monthly returns

myStocks_returns_monthly_summary %>%
  
  kable() %>% 
  
  kable_styling()
  
```

We now create a density plot, using `geom_density()`, for each of the stocks to get a better sense of the distribution of the returns per stock.

```{r density_monthly_returns}

myStocks_returns_monthly %>%
  
  # only display density of the stocks
  
  filter(symbol != "SPY") %>%
  
  # add some color
  
  ggplot(aes(x = monthly_returns, fill = symbol)) +
  
  geom_density() + 
  
   # create a plot for each stock
  
  facet_wrap(~symbol) +
  
  labs(title = "Distribution of monthly returns per stock", 
       x = "Monthly returns",
       y = "Density") +
  
  # hide legend
  
  theme(legend.position = "none") 

```

>From this plot, we can infer that Blackrock (BLK) is the riskiest stock with the highest volatility given the broad distribution / deviation from 0.0 of BLK returns. On the other hand, Coca Cola (KO) is the least risky as the plot for Coca Cola displays a very low deviation of returns from 0.0.

Finally, we plot the relationship between mean expected monthly returns and the return's standard deviation. 

```{r risk_return_plot}

myStocks_returns_monthly_summary %>% 
  
  # only display the stocks
  
  filter(symbol != "SPY") %>% 
  
  # plot scatterplot and increase size of points
  
  ggplot(aes(x=sd, y = mean,
             label = symbol)) +
  
  geom_point(aes(colour = symbol),
             size = 4) +
  
  # add labels to the points
  
  ggrepel::geom_text_repel(aes(colour = symbol)) + 
  
  # add title and labels to graph
  
  labs(title = 'Do higher risks yield higher returns?',
       subtitle = 'Risk/Return Profile of selected Stocks',
       x = 'Risk (SD of monthly returns)', 
       y ="Average monthly return") +
  
  # use a nice theme
  
  theme_bw() + 
  
  # display x values as percentages
  
  scale_x_continuous(labels = scales::percent) + 
  
  # display y values as percentages
  
  scale_y_continuous(labels = scales::percent) + 
  
  # hide legend
  
  theme(legend.position = "none") 

```

> The plot shows Amazon (AMZN) is the stock with the strongest risk-reward profile: It yields the highest average monthly returns and is the riskiest of the selected stocks as indicated by its position on the top right corner of the graph. Indeed, higher risks generally lead to a higher mean monthly return. However, BP's position proves that a high risk does not necessarily imply a high return - BP has the lowest monthly returns while the BP stock is the second riskiest. 
